---
title: "Activity Classification Using Personal Activity Devices"
author: "Jason LaJeunesse"
date: "Sunday, November 22, 2015"
output: html_document
---

###Overview

The purpose of this assignment is to use data from personal activity devices to classify activities as opposed to simply aggregating usage data.So this is a look at the quality of the activity data as opposed to just quantity using machine learning

###Analysis

The first step is to load the associated libraries and test data set. Then data is separated into a training and validation set.

```{r}

library(AppliedPredictiveModeling)
library(caret)
library(sqldf)
library(gbm)

training <- read.csv("E:/Coursera/machine_learning/pml-training.csv")
set.seed(1337)

# Pre-Processing

# Delete columns that have any NA values
training <- training[, !is.na(colSums( training == 'NA'))]

# Separate out 25% of the data for validation purposes at the end
training_partition = createDataPartition(training$classe, p = 0.75)[[1]]
training_set = training[training_partition,]
testing_set = training[-training_partition,]
```

For this analysis, a boosted tree model with cross validation will be used to create a classification model. The cross validation uses 5 folds and 2 repeats. While this model could be more efficient, almost of the columns that are not an average, max, min, standard deviation are used to increase model accuracy. Predictors like window number do not make sense and have been excluded.

```{r}

# 5 fold cross validation control
control <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

# For this model, I've chosen to use a quick and dirty select of all values that are not
# averages, maxes, mins, or standard deviations. This uses the boosted tree model "gbm" package
train_model1 <- train(training_set$classe ~ 
                        user_name  +
                        roll_belt	+
                        pitch_belt	+
                        yaw_belt	+
                        total_accel_belt	+
                        gyros_belt_x	+
                        gyros_belt_y	+
                        gyros_belt_z	+
                        accel_belt_x	+
                        accel_belt_y	+
                        accel_belt_z	+
                        magnet_belt_x	+
                        magnet_belt_y	+
                        magnet_belt_z	+
                        roll_arm	+
                        pitch_arm	+
                        yaw_arm	+
                        total_accel_arm	+
                        gyros_arm_x	+
                        gyros_arm_y	+
                        gyros_arm_z	+
                        accel_arm_x	+
                        accel_arm_y	+
                        accel_arm_z	+
                        magnet_arm_x	+
                        magnet_arm_y	+
                        magnet_arm_z	+
                        roll_dumbbell	+
                        pitch_dumbbell	+
                        yaw_dumbbell	+
                        total_accel_dumbbell	+
                        gyros_dumbbell_x	+
                        gyros_dumbbell_y	+
                        gyros_dumbbell_z	+
                        accel_dumbbell_x	+
                        accel_dumbbell_y	+
                        accel_dumbbell_z	+
                        magnet_dumbbell_x	+
                        magnet_dumbbell_y	+
                        magnet_dumbbell_z	+
                        roll_forearm	+
                        pitch_forearm	+
                        yaw_forearm	+
                        total_accel_forearm	+
                        gyros_forearm_x	+
                        gyros_forearm_y	+
                        gyros_forearm_z	+
                        accel_forearm_x	+
                        accel_forearm_y	+
                        accel_forearm_z	+
                        magnet_forearm_x	+
                        magnet_forearm_y	+
                        magnet_forearm_z
                      ,data = training_set, trControl = control, method = "gbm")

# Confusion matrix using the testing_set created initially with 25% of the test data
confusionMatrix(predict(train_model1, testing_set), testing_set$classe)
```
**Figure 1:** Confusion Matrix against Training Set

From the first figure, it can be seen that the model which used cross-validation is over 95% accurate. This would mean the error rate should be less than 5% for samples from the same set. This was validated on the "pml-testing.csv" with an error rate of 0%. 

```{r}
# Predict the 20 classifications from the pml-testing file
testing <- read.csv("E:/Coursera/machine_learning/pml-testing.csv")
answers = predict(train_model1, testing)
```

